---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

This is the R script used for phenotype validation of ML model accompanying the following paper
Title: AutMedAI: Predicting Autism from a Minimal Set of Medical and Background Information using Machine Learning
Authors: Shyam Sundar Rajagopalan, Yali Zhang, Ashraf Yahia, and Kristiina Tammimies
Correspondence: Dr. Kristiina Tammimies, kristiina.tammimies@ki.se

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Score distribution

```{r,message=TRUE}
library(readxl)
library(reshape2)
library(ggplot2)
library(ggpubr)

validation_data_v10 <- read_excel('v10_validation_data.xlsx',sheet = 1)

# ASD
ASD_df <- as.data.frame(validation_data_v10[validation_data_v10$asd_x == 1,]) # 10476 asd

# label the prediction result
for (i in 1:nrow(ASD_df)) {
  if (ASD_df$asd_x[i] == ASD_df$predicted_label[i]){
    ASD_df[i,'Result'] <- 'Yes'
  }else{
    ASD_df[i,'Result'] <- 'No'
  }
}
table(ASD_df$Result) # No 2214, Yes 8262

# extract variables and annotate to individuals in dataset
var_list_v10 <- read.table('phenotype_compare_selection_v10.txt',sep = '\t',comment.char = "#",header = T)
var_list_v10

var_name_list <- c()
for (i in 1:nrow(var_list_v10)){
  var_name <- paste0(var_list_v10$phenotype[i],':',var_list_v10$variable[i])
  var_name_list <- c(var_name_list,var_name)
  var <- var_list_v10$variable[i]
  file_route <- var_list_v10$file[i]
  
  filein <- read.csv(paste0('../../phenotype_dataset/SPARK_collection_v10_2023-07-21/',file_route),header = T)
  
  for (j in 1:length(ASD_df$subject_sp_id)){
    if (ASD_df$subject_sp_id[j] %in% filein$subject_sp_id){
      input <- filein[filein$subject_sp_id == ASD_df$subject_sp_id[j],var]
      if (length(input) > 1){
        for (k in 1:length(input)){
          if (k == 1){
            input_1 <- filein[filein$subject_sp_id == ASD_df$subject_sp_id[j],var][k]
            year_1 <- filein[filein$subject_sp_id == ASD_df$subject_sp_id[j],'test_year'][k]
          }else{
            input_2 <- filein[filein$subject_sp_id == ASD_df$subject_sp_id[j],var][k]
            year_2 <- filein[filein$subject_sp_id == ASD_df$subject_sp_id[j],'test_year'][k]
            if (year_2 > year_1){
              input_1 <- input_2
              year_1 <- year_2
            }
          }
        }
        input <- input_1
      }
      ASD_df[j,var_name] <- input
    }else{
      ASD_df[j,var_name] <- 'pheno_missing'
    }
  }
}

# variable recording summary
var_subj_summary_ASD <- data.frame(row.names = var_name_list)
for (i in var_name_list){
  na_n <- sum(is.na(ASD_df[,i]))
  miss_n <- length(which(ASD_df[,i] == 'pheno_missing'))
  value_n <- sum(table(ASD_df[,i])) - length(which(ASD_df[,i] == 'pheno_missing'))
  var_subj_summary_ASD[i,'Record'] <- value_n
  var_subj_summary_ASD[i,'NA'] <- na_n
  var_subj_summary_ASD[i,'Missing'] <- miss_n
  var_subj_summary_ASD[i,'Record_percentage'] <- value_n/length(ASD_df$subject_sp_id)
}

variable_data_ASD <- ASD_df[,c(var_name_list,'Result')]
str(variable_data_ASD)
```

```{r,warning=FALSE}
for (i in var_name_list) {
  plot_df <- variable_data_ASD[-which(is.na(variable_data_ASD[,i])),]
  plot_df <- plot_df[-which(plot_df[,i] == 'pheno_missing'),]
  if (length(plot_df[plot_df$Result == 'Yes',i] )< 5000){
    a <- shapiro.test(as.numeric(plot_df[plot_df$Result == 'Yes',i]))
    print(paste0(i,'\npredict correctly: ',a[["p.value"]]))
  }else{
    print(paste0(i,'\npredict correctly: '))
    qqnorm(as.numeric(plot_df[plot_df$Result == 'Yes',i]))
    qqline(as.numeric(plot_df[plot_df$Result == 'Yes',i]))
  }
  if (length(plot_df[plot_df$Result == 'No',i] )< 5000){
    b <- shapiro.test(as.numeric(plot_df[plot_df$Result == 'No',i]))
    print(paste0(i,'\npredict wrongly: ',b[["p.value"]]))
  }else{
    qqnorm(as.numeric(plot_df[plot_df$Result == 'Yes',i]))
    qqline(as.numeric(plot_df[plot_df$Result == 'Yes',i]))
  }
}
```

```{r}
# Plot data distribution
for (i in var_name_list){
  
  modified_string <- gsub("_", " ", i)
  modified_string <- gsub(":", ": ", modified_string)
  modified_string <- tools::toTitleCase(modified_string)
  
  plot_df <- variable_data_ASD[-which(is.na(variable_data_ASD[,i])),]
  plot_df <- plot_df[-which(plot_df[,i] == 'pheno_missing'),]

  ggplot(plot_df,aes(x=Result,y=as.numeric(plot_df[,i]),fill = Result))+
    geom_jitter(aes(color=Result),position = position_jitter(0.2),size=0.5,alpha=0.6)+
    geom_boxplot(alpha=1,outlier.size = 0.5,size = 0.5,width=0.2,fill='transparent')+
    geom_violin(alpha=0.5)+
    xlab('')+
    ylab(modified_string)+
    ylim(min(as.numeric(plot_df[,i]))-1,max(as.numeric(plot_df[,i]))+max(as.numeric(plot_df[,i]))/10)+
    theme_bw()+
    geom_signif(comparisons = list(c("No", "Yes")),
                map_signif_level = T,
                test = wilcox.test,
                textsize=3)+
    theme(legend.position = 'none',
          plot.title = element_text(size=10))+
    scale_x_discrete(breaks = c('No','Yes'),
                     labels = c(paste0(c('No','Yes'),'\n(n=',c(sum(plot_df$Result == 'No'),sum(plot_df$Result == 'Yes')),')')))
  # ggsave(paste0(i,'_violine.png'),plot = last_plot(),width = 2.5,height = 3.8,dpi = 600,limitsize = TRUE)

  var_subj_summary_ASD[i,'Predict_correct'] <- length(plot_df[plot_df$Result == 'Yes',i])
  var_subj_summary_ASD[i,'Predict_wrong'] <- length(plot_df[plot_df$Result == 'No',i])
  wilcox_test <- wilcox.test(as.numeric(plot_df[plot_df$Result == 'Yes',i]), as.numeric(plot_df[plot_df$Result == 'No',i]))
  var_subj_summary_ASD[i,'wilcox_test_p.value'] <- wilcox_test$p.value
}
var_subj_summary_ASD
```

```{r,warning=FALSE}

# Put on one figure
melt_df_ASD <- melt(variable_data_ASD,id.vars="Result")
melt_df_ASD <- melt_df_ASD[-which(is.na(melt_df_ASD$value)),]
melt_df_ASD <- melt_df_ASD[-which(melt_df_ASD$value == 'pheno_missing'),]
melt_df_ASD$value <- as.numeric(melt_df_ASD$value)
min(melt_df_ASD$value)

melt_df_ASD$variable <- as.character(melt_df_ASD$variable)

for (i in var_name_list){
  modified_string <- gsub("_", " ", i)
  modified_string <- gsub(":", ": ", modified_string)
  modified_string <- tools::toTitleCase(modified_string)
  melt_df_ASD[,'variable'][which(melt_df_ASD[,'variable']==i)] <- modified_string
}

# unique(melt_df_ASD$variable)

plot_index <- 0 
label_set <- function(x) {
  plot_index <<- plot_index + 1L
  switch(plot_index,
         c(paste0(c('No','Yes'),'\n(n=',c(sum(melt_df_ASD$Result == 'No' & melt_df_ASD$variable=="CBCL1-5: Total Problems t Score"),sum(melt_df_ASD$Result == 'Yes' & melt_df_ASD$variable=="CBCL1-5: Total Problems t Score")),')')),
         c(paste0(c('No','Yes'),'\n(n=',c(sum(melt_df_ASD$Result == 'No' & melt_df_ASD$variable=="CBCL6-18: Total Problems t Score"),sum(melt_df_ASD$Result == 'Yes' & melt_df_ASD$variable=="CBCL6-18: Total Problems t Score")),')')),
         c(paste0(c('No','Yes'),'\n(n=',c(sum(melt_df_ASD$Result == 'No' & melt_df_ASD$variable=="IQ: Fsiq Score"),sum(melt_df_ASD$Result == 'Yes' & melt_df_ASD$variable=="IQ: Fsiq Score")),')')),
         c(paste0(c('No','Yes'),'\n(n=',c(sum(melt_df_ASD$Result == 'No' & melt_df_ASD$variable=="Scq: Final Score" ),sum(melt_df_ASD$Result == 'Yes' & melt_df_ASD$variable=="Scq: Final Score" )),')')))
}

ggplot(melt_df_ASD,aes(x=Result,y=value,fill = Result))+
  geom_jitter(aes(color=Result),position = position_jitter(0.2),size=0.5,alpha=0.6)+
  geom_boxplot(alpha=1,outlier.size = 0.5,size = 0.5,width=0.2,fill='transparent')+
  geom_violin(alpha=0.5)+
  facet_wrap(~variable,nrow = 1,scales = 'free') +
  xlab('')+
  ylab('Score')+
  geom_signif(comparisons = list(c("No", "Yes")),
              map_signif_level = T,
              test = wilcox.test,
              textsize=3)+
  theme_bw()+
  theme(legend.position = 'none',
        strip.background=element_rect(colour="black",fill="white"))+
  scale_x_discrete(labels = label_set)
# ggsave(paste0('total','_violine.pdf'),plot = last_plot(),width = 9.5,height = 3.8,dpi = 600)
```

```{r,warning=FALSE}

## Non-ASD
non_ASD_df <- as.data.frame(validation_data_v10[validation_data_v10$asd_x == 0,]) # 10476 asd
table(non_ASD_df$asd_x) # 1460

# label the prediction result
for (i in 1:nrow(non_ASD_df)) {
  if (non_ASD_df$asd_x[i] == non_ASD_df$predicted_label[i]){
    non_ASD_df[i,'Result'] <- 'Yes'
  }else{
    non_ASD_df[i,'Result'] <- 'No'
  }
}
table(non_ASD_df$Result) # No 305, Yes 1155

var_name_list <- c()
for (i in 1:nrow(var_list_v10)){
  var_name <- paste0(var_list_v10$phenotype[i],':',var_list_v10$variable[i])
  var_name_list <- c(var_name_list,var_name)
  var <- var_list_v10$variable[i]
  file_route <- var_list_v10$file[i]
  
  filein <- read.csv(paste0('../../phenotype_dataset/SPARK_collection_v10_2023-07-21/',file_route),header = T)
  
  for (j in 1:length(non_ASD_df$subject_sp_id)){
    if (non_ASD_df$subject_sp_id[j] %in% filein$subject_sp_id){
      input <- filein[filein$subject_sp_id == non_ASD_df$subject_sp_id[j],var]
      if (length(input) > 1){
        for (k in 1:length(input)){
          if (k == 1){
            input_1 <- filein[filein$subject_sp_id == non_ASD_df$subject_sp_id[j],var][k]
            year_1 <- filein[filein$subject_sp_id == non_ASD_df$subject_sp_id[j],'test_year'][k]
          }else{
            input_2 <- filein[filein$subject_sp_id == non_ASD_df$subject_sp_id[j],var][k]
            year_2 <- filein[filein$subject_sp_id == non_ASD_df$subject_sp_id[j],'test_year'][k]
            if (year_2 > year_1){
              input_1 <- input_2
              year_1 <- year_2
            }
          }
        }
        input <- input_1
      }
      non_ASD_df[j,var_name] <- input
    }else{
      non_ASD_df[j,var_name] <- 'pheno_missing'
    }
  }
}

# variable recording summary
var_subj_summary_non_ASD <- data.frame(row.names = var_name_list)
for (i in var_name_list){
  na_n <- sum(is.na(non_ASD_df[,i]))
  miss_n <- length(which(non_ASD_df[,i] == 'pheno_missing'))
  value_n <- sum(table(non_ASD_df[,i])) - length(which(non_ASD_df[,i] == 'pheno_missing'))
  var_subj_summary_non_ASD[i,'Record'] <- value_n
  var_subj_summary_non_ASD[i,'NA'] <- na_n
  var_subj_summary_non_ASD[i,'Missing'] <- miss_n
  var_subj_summary_non_ASD[i,'Record_percentage'] <- value_n/length(non_ASD_df$subject_sp_id)
}


modified_string <- gsub("_", " ", 'scq:final_score')
modified_string <- gsub(":", ": ", modified_string)
modified_string <- tools::toTitleCase(modified_string)

plot_df <- non_ASD_df[-which(is.na(non_ASD_df[,'scq:final_score'])),]
plot_df <- plot_df[-which(plot_df[,'scq:final_score'] == 'pheno_missing'),]

ggplot(plot_df,aes(x=Result,y=as.numeric(plot_df[,'scq:final_score']),fill = Result))+
  geom_jitter(aes(color=Result),position = position_jitter(0.2),size=0.5,alpha=0.6)+
  geom_boxplot(alpha=1,outlier.size = 0.5,size = 0.5,width=0.15,fill='transparent')+
  geom_violin(alpha=0.5)+
  xlab('')+
  ylab(modified_string)+
  ylim(-1,max(as.numeric(plot_df[,i]))+max(as.numeric(plot_df[,i]))/10)+
  theme_bw()+
  geom_signif(comparisons = list(c("No", "Yes")),
              map_signif_level = T,
              test = wilcox.test,
              textsize=3)+
  theme(legend.position = 'none',
        plot.title = element_text(size=10))+
  scale_x_discrete(breaks = c('No','Yes'),
                   labels = c(paste0(c('No','Yes'),'\n(n=',c(sum(plot_df$Result == 'No'),sum(plot_df$Result == 'Yes')),')')))
# ggsave(paste0('False_positive/',i,'_violine.png'),plot = last_plot(),width = 2.5,height = 3.8,dpi = 600,limitsize = TRUE)

```

```{r}
var_subj_summary_non_ASD['scq:final_score','Result_Ture'] <- length(plot_df[plot_df$Result == 'Yes','scq:final_score'])
var_subj_summary_non_ASD['scq:final_score','Result_False'] <- length(plot_df[plot_df$Result == 'No','scq:final_score'])

shapiro.test(as.numeric(plot_df[plot_df$Result == 'Yes','scq:final_score']))
```

```{r}
shapiro.test(as.numeric(plot_df[plot_df$Result == 'No','scq:final_score']))
```

```{r}
wilcox_test <- wilcox.test(as.numeric(plot_df[plot_df$Result == 'Yes','scq:final_score']), as.numeric(plot_df[plot_df$Result == 'No','scq:final_score']))
var_subj_summary_non_ASD['scq:final_score','wilcox_test_p.value'] <- wilcox_test$p.value
var_subj_summary_non_ASD
```

## Odds ratio

```{r,message=FALSE,error=TRUE,warning=FALSE}
v_file <- as.data.frame(read_excel('pheno_selection.xlsx',sheet = 1, na = 'NA'))
BMS_V10 <- read.csv("../../phenotype_dataset/SPARK_collection_v10_2023-07-21/basic_medical_screening_production-2023-07-21.csv",header = T)

# asd
for (i in 1:length(v_file$variable)){
  var <- v_file$variable[i]
  for (j in 1:length(ASD_df$subject_sp_id)){
    if (ASD_df$subject_sp_id[j] %in% BMS_V10$subject_sp_id){
      input <- BMS_V10[BMS_V10$subject_sp_id == ASD_df$subject_sp_id[j],var]
      if (length(input) > 1){
        for (k in 1:length(input)){
          if (k == 1){
            input_1 <- BMS_V10[BMS_V10$subject_sp_id == ASD_df$subject_sp_id[j],var][k]
            year_1 <- BMS_V10[BMS_V10$subject_sp_id == ASD_df$subject_sp_id[j],'test_year'][k]
          }else{
            input_2 <- BMS_V10[BMS_V10$subject_sp_id == ASD_df$subject_sp_id[j],var][k]
            year_2 <- BMS_V10[BMS_V10$subject_sp_id == ASD_df$subject_sp_id[j],'test_year'][k]
            if (year_2 > year_1){
              input_1 <- input_2
              year_1 <- year_2
            }
          }
        }
        input <- input_1
      }
      if(is.na(input)){
        ASD_df[j,var] <- 'blank'
      }else{
        ASD_df[j,var] <- input
      }
    }else{
      ASD_df[j,var] <- 'pheno_missing'
    }
  }
}

var_subj_summary_ASD_pheno2 <- data.frame(row.names = v_file$variable)
for (i in v_file$variable){
  na_n <- length(which(ASD_df[,i] == 'blank'))
  miss_n <- length(which(ASD_df[,i] == 'pheno_missing'))
  value_n <- sum(table(ASD_df[,i])) - length(which(ASD_df[,i] == 'pheno_missing'))
  
  var_subj_summary_ASD_pheno2[i,'Record'] <- value_n
  var_subj_summary_ASD_pheno2[i,'NA'] <- na_n
  var_subj_summary_ASD_pheno2[i,'Missing'] <- miss_n
  var_subj_summary_ASD_pheno2[i,'Record_percentage'] <- value_n/length(ASD_df$subject_sp_id)
}
var_subj_summary_ASD_pheno2
```

```{r,warning=FALSE}
# non-asd
for (i in 1:length(v_file$variable)){
  var <- v_file$variable[i]
  for (j in 1:length(non_ASD_df$subject_sp_id)){
    if (non_ASD_df$subject_sp_id[j] %in% BMS_V10$subject_sp_id){
      input <- BMS_V10[BMS_V10$subject_sp_id == non_ASD_df$subject_sp_id[j],var]
      if (length(input) > 1){
        for (k in 1:length(input)){
          if (k == 1){
            input_1 <- BMS_V10[BMS_V10$subject_sp_id == non_ASD_df$subject_sp_id[j],var][k]
            year_1 <- BMS_V10[BMS_V10$subject_sp_id == non_ASD_df$subject_sp_id[j],'test_year'][k]
          }else{
            input_2 <- BMS_V10[BMS_V10$subject_sp_id == non_ASD_df$subject_sp_id[j],var][k]
            year_2 <- BMS_V10[BMS_V10$subject_sp_id == non_ASD_df$subject_sp_id[j],'test_year'][k]
            if (year_2 > year_1){
              input_1 <- input_2
              year_1 <- year_2
            }
          }
        }
        input <- input_1
      }
      if(is.na(input)){
        non_ASD_df[j,var] <- 'blank'
      }else{
        non_ASD_df[j,var] <- input
      }
    }else{
      non_ASD_df[j,var] <- 'pheno_missing'
    }
  }
}

var_subj_summary_non_ASD_pheno2 <- data.frame(row.names = v_file$variable)
for (i in v_file$variable){
  na_n <- length(which(non_ASD_df[,i] == 'blank'))
  miss_n <- length(which(non_ASD_df[,i] == 'pheno_missing'))
  value_n <- sum(table(non_ASD_df[,i])) - length(which(non_ASD_df[,i] == 'pheno_missing'))
  
  var_subj_summary_non_ASD_pheno2[i,'Record'] <- value_n
  var_subj_summary_non_ASD_pheno2[i,'NA'] <- na_n
  var_subj_summary_non_ASD_pheno2[i,'Missing'] <- miss_n
  var_subj_summary_non_ASD_pheno2[i,'Record_percentage'] <- value_n/length(non_ASD_df$subject_sp_id)
}
var_subj_summary_non_ASD_pheno2
```

```{r,warning=FALSE}

validation_data_merge <- rbind.data.frame(cbind.data.frame(ASD_df[,c('Result',v_file$variable)],Outcome=rep('ASD',nrow(ASD_df))),
                                          cbind.data.frame(non_ASD_df[,c('Result',v_file$variable)],Outcome=rep('Non-ASD',nrow(non_ASD_df))))

for (i in v_file$variable){
  validation_data_merge[which(validation_data_merge[,i]=='blank'),i] <- 0
}

# summary
df_version <- c()
df_predict <- c()
df_phe <- c()
df_nall <- c()
df_nhave <- c()
df_freq <- c()
for (i in c('Yes','No')) {
  for (v in c('ASD','Non-ASD')){
    for (j in v_file$variable) {
      df_version <- c(df_version,v)
      df_predict <- c(df_predict,i)
      df_phe <- c(df_phe,j)
      df_nhave <- c(df_nhave,sum(validation_data_merge[validation_data_merge$Result == i & validation_data_merge$Outcome==v ,j] > 0))
      df_nall <- c(df_nall,sum(validation_data_merge$Result == i  & validation_data_merge$Outcome==v))
      df_freq <- c(df_freq,sum(validation_data_merge[validation_data_merge$Result == i  & validation_data_merge$Outcome==v,j] > 0)/sum(validation_data_merge$Result == i  & validation_data_merge$Outcome==v))
    }
  }
}

freq_df <- cbind.data.frame('Outcome' = df_version,
                            'Result' = df_predict,
                            'Phenotype' = df_phe,
                            'Num_with' = df_nhave,
                            'Num_all' = df_nall,
                            'Freq' = df_freq)

# Odds ratio
library(epitools)

for (i in unique(validation_data_merge$Outcome)){
  result_list <- validation_data_merge[validation_data_merge$Outcome == i ,'Result']
  for (phe in v_file$variable){
    diagnosis <- validation_data_merge[validation_data_merge$Outcome == i ,phe]
    table <- table(result_list, diagnosis)
    freq_df[freq_df$Phenotype == phe & freq_df$Outcome == i ,'OR'] <- oddsratio(table)$measure[2,1]
    freq_df[freq_df$Phenotype == phe & freq_df$Outcome == i ,'lower_CI'] <- oddsratio(table)$measure[2,2]
    freq_df[freq_df$Phenotype == phe & freq_df$Outcome == i ,'upper_CI'] <- oddsratio(table)$measure[2,3]
  }
}

for ( i in 1:nrow(freq_df)){
  type <- v_file[v_file$variable == freq_df$Phenotype[i],'variable_group']
  freq_df[i,'Type'] <- tools::toTitleCase(type)
}

freq_df$Phenotype <- as.factor(freq_df$Phenotype)
freq_df$Phenotype <- factor(freq_df$Phenotype,levels = c('Attention or behavior disorders','Attention Deficit-Hyperactivity Disorder',
                                                         'Conduct Disorder','Intermittent Explosive Disorder','Oppositional Defiant Disorder',
                                                         'Language delay or language disorder',  'Learning disability', 'Motor delay or developmental coordination disorder',
                                                         'Mutism', 'Social (Pragmatic) Communication Disorder' ,        
                                                         'Speech articulation problems') )

## forest plot
df_forest_ASD <- data.frame(row.names = levels(freq_df$Phenotype))

for (i in 1:nlevels(freq_df$Phenotype)) {
  phe <- levels(freq_df$Phenotype)[i]
  df_forest_ASD[i,'Diagnosis'] <- phe
  df_forest_ASD[i,'Yes_event'] <-  freq_df[which(freq_df$Phenotype == phe & freq_df$Outcome=='ASD' & freq_df$Result=='Yes'),'Num_with']
  df_forest_ASD[i,'Yes_all'] <-  freq_df[freq_df$Phenotype == phe & freq_df$Outcome=='ASD' & freq_df$Result=='Yes','Num_all']
  df_forest_ASD[i,'No_event'] <-  freq_df[freq_df$Phenotype == phe & freq_df$Outcome=='ASD' & freq_df$Result=='No','Num_with']
  df_forest_ASD[i,'No_all'] <-  freq_df[freq_df$Phenotype == phe & freq_df$Outcome=='ASD' & freq_df$Result=='No','Num_all']
  df_forest_ASD[i,'OR'] <-  freq_df[freq_df$Phenotype == phe & freq_df$Outcome=='ASD' & freq_df$Result=='No','OR']  
  df_forest_ASD[i,'lower_CI'] <-  freq_df[freq_df$Phenotype == phe & freq_df$Outcome=='ASD' & freq_df$Result=='No','lower_CI']
  df_forest_ASD[i,'upper_CI'] <-  freq_df[freq_df$Phenotype == phe & freq_df$Outcome=='ASD' & freq_df$Result=='No','upper_CI']
  df_forest_ASD[i,'OR (95% CI)'] <-  paste0(sprintf("%.2f", as.numeric(df_forest_ASD[i,'OR'])),' (',sprintf("%.2f", df_forest_ASD[i,'lower_CI']),'-',sprintf("%.2f", df_forest_ASD[i,'upper_CI']),')  ')
}

df_forest_non_ASD <- data.frame(row.names = levels(freq_df$Phenotype))

for (i in 1:nlevels(freq_df$Phenotype)) {
  phe <- levels(freq_df$Phenotype)[i]
  df_forest_non_ASD[i,'Diagnosis'] <- phe
  df_forest_non_ASD[i,'Yes_event'] <-  freq_df[which(freq_df$Phenotype == phe & freq_df$Outcome=='Non-ASD' & freq_df$Result=='Yes'),'Num_with']
  df_forest_non_ASD[i,'Yes_all'] <-  freq_df[freq_df$Phenotype == phe & freq_df$Outcome=='Non-ASD' & freq_df$Result=='Yes','Num_all']
  df_forest_non_ASD[i,'No_event'] <-  freq_df[freq_df$Phenotype == phe & freq_df$Outcome=='Non-ASD' & freq_df$Result=='No','Num_with']
  df_forest_non_ASD[i,'No_all'] <-  freq_df[freq_df$Phenotype == phe & freq_df$Outcome=='Non-ASD' & freq_df$Result=='No','Num_all']
  df_forest_non_ASD[i,'OR'] <-  freq_df[freq_df$Phenotype == phe & freq_df$Outcome=='Non-ASD' & freq_df$Result=='No','OR']  
  df_forest_non_ASD[i,'lower_CI'] <-  freq_df[freq_df$Phenotype == phe & freq_df$Outcome=='Non-ASD' & freq_df$Result=='No','lower_CI']
  df_forest_non_ASD[i,'upper_CI'] <-  freq_df[freq_df$Phenotype == phe & freq_df$Outcome=='Non-ASD' & freq_df$Result=='No','upper_CI']
  df_forest_non_ASD[i,'OR (95% CI)'] <-  paste0(sprintf("%.2f", as.numeric(df_forest_non_ASD[i,'OR'])),' (',sprintf("%.2f", df_forest_non_ASD[i,'lower_CI']),'-',sprintf("%.2f", df_forest_non_ASD[i,'upper_CI']),')  ')
}

test_df <- cbind(df_forest_ASD[,c(1,2,4,9)],` `=paste(rep(" ", nrow(df_forest_ASD)), collapse = "    "),df_forest_ASD[,c(6,7,8)])
test_df_2 <- cbind(df_forest_non_ASD[,c(1,2,4,9)],` `=paste(rep(" ", nrow(df_forest_non_ASD)), collapse = "    "),df_forest_non_ASD[,c(6,7,8)])

row1 <- c('ASD','Yes\n(n=8268)','No\n(n=2214)    ','','','','','')
row2 <- c('Non-ASD','Yes\n(n=1155)','No\n(n=305)','','','','','')

forest_df <- rbind(row1,test_df,row2,test_df_2)
forest_df2 <- forest_df[1:21,]

forest_df2$OR <- round(as.numeric(forest_df2$OR),2)

forest_df2$OR[1] <- ''
forest_df2$OR[13] <- ''
colnames(forest_df2)[2] <- 'AutMedAI prediction\nlabel correct'
colnames(forest_df2)[3] <- ' '
colnames(forest_df2)[5] <- '  '

forest_df2$Diagnosis <- paste0('   ',forest_df2$Diagnosis)
forest_df2$Diagnosis[1] <- 'Autism'
forest_df2$Diagnosis[13] <- 'Non-autism'

tm <- forest_theme(base_size = 10,
                   # Confidence interval point shape, line type/color/width
                   ci_pch = 15,
                   ci_col = "black",
                   ci_fill = "darkblue",
                   ci_alpha = 0.8,
                   ci_lty = 1,
                   ci_lwd = 1.5,
                   ci_Theight = 0.2, # Set a T end at the end of CI 
                   # Reference line width/type/color
                   refline_lwd = gpar(lwd = 1, lty = "dashed", col = "grey"),
                   core=list(bg_params=list(fill = c("grey95",rep("White",11) ))))

p = forest(forest_df2[,1:5],
       est = as.numeric(forest_df2$OR),
       lower = as.numeric(forest_df2$lower_CI), 
       upper = as.numeric(forest_df2$upper_CI),
       # sizes = dt$se,
       ci_column = 5,
       ref_line = 1,
       xlim = c(-1, 6),
       xlab = c("            OR (95% CI)"),
       theme = tm)
# ggsave('forest_240529.pdf',plot = p,width = 10,height = 6,dpi = 600,limitsize = TRUE)
```

```{r}

# adjust p value

freq_df$No_n <- freq_df$Num_all - freq_df$Num_with

test_df2 <- data.frame(row.names = c(0,1))

for (v in unique(freq_df$Outcome)){
  for (i in unique(freq_df$Phenotype)){
    print(c(v,i))
    test_df2[1,'Yes'] <- freq_df[freq_df$Phenotype == i & freq_df$Result == 'Yes'  & freq_df$Outcome == v, 'No_n']
    test_df2[2,'Yes'] <- freq_df[freq_df$Phenotype == i & freq_df$Result == 'Yes' & freq_df$Outcome == v, 'Num_with']
    test_df2[1,'No'] <- freq_df[freq_df$Phenotype == i & freq_df$Result == 'No' & freq_df$Outcome == v, 'No_n']
    test_df2[2,'No'] <- freq_df[freq_df$Phenotype == i & freq_df$Result == 'No' & freq_df$Outcome == v, 'Num_with']
    
    test_df2$Yes <- as.numeric(test_df2$Yes)
    test_df2$No <- as.numeric(test_df2$No)
    
    chisq <- chisq.test(test_df2)
    
    freq_df[freq_df$Phenotype == i & freq_df$Outcome == v,'Chisq_p'] <- chisq$p.value
    adj_p <- chisq$p.value * length(unique(freq_df$Phenotype))
    freq_df[freq_df$Phenotype == i & freq_df$Outcome == v,'Chisq_p_adjust'] <- adj_p
   
     if (adj_p <= 0.001){
      freq_df[freq_df$Phenotype == i & freq_df$Outcome == v,'significance'] <- '***'
    }else if(adj_p > 0.001 & adj_p <= 0.01){
      freq_df[freq_df$Phenotype == i & freq_df$Outcome == v,'significance'] <- '**'
    }else if(adj_p > 0.01 & adj_p <= 0.05){
      freq_df[freq_df$Phenotype == i & freq_df$Outcome == v,'significance'] <- '*'
    }else if(adj_p > 0.05){
      freq_df[freq_df$Phenotype == i & freq_df$Outcome == v,'significance'] <- 'NA'
    }
  }
}

# output data table

library(writexl)
var_subj_summary_ASD2 <- cbind('phenotype' = rownames(var_subj_summary_ASD),var_subj_summary_ASD)
var_subj_summary_non_ASD2 <- cbind('phenotype' = rownames(var_subj_summary_non_ASD),var_subj_summary_non_ASD)
# write_xlsx(list('ASD_score'=var_subj_summary_ASD2, 'Non-ASD_score'=var_subj_summary_non_ASD2,"Odds ratio"=freq_df),"statistics_model_prediction.xlsx")

freq_df
```
